---
title: "Practical Machine Learning"
output:
  html_document:
    keep_md: yes
    toc: yes
---

```{r echo=FALSE, message=FALSE}
setwd("C:/Users/Administrator/Git/DataScienceCoursera/Practical Machine Learning")
```

## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data
```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

### Downloading
```{r, cache = T}
srcTrainingData <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
srcTestingData <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
fTrainingData <- "./data/pml-training.csv"
fTestingData  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(fTrainingData)) {
  download.file(srcTrainingData, destfile=fTrainingData)
}
if (!file.exists(fTestingData)) {
  download.file(srcTestingData, destfile=fTestingData)
}
```  
### Reading
```{r, cache = T}
TrainingData <- read.csv("./data/pml-training.csv")
TestingData <- read.csv("./data/pml-testing.csv")
dim(TrainingData)
dim(TestingData)
```

Variable CLASSE in the training data will be predicted.

### Preprocessing
```{r, cache = T}
sum(complete.cases(TrainingData))
```


```{r, cache = T}
TrainingData <- TrainingData[, colSums(is.na(TrainingData)) == 0] 
TestingData <- TestingData[, colSums(is.na(TestingData)) == 0] 
```  

```{r, cache = T}
classe <- TrainingData$classe
trainingProcessed1 <- grepl("^X|timestamp|window", names(TrainingData))
TrainingData <- TrainingData[, !trainingProcessed1]
trainingProcessed2 <- TrainingData[, sapply(TrainingData, is.numeric)]
trainingProcessed2$classe <- classe
testingProcessed <- grepl("^X|timestamp|window", names(TestingData))
TestingData <- TestingData[, !testingProcessed]
testingProcessed2 <- TestingData[, sapply(TestingData, is.numeric)]
```

### Separation
Data will be seperated into 70% for Training and 30% for Validation

```{r, cache = T}
set.seed(4242) # For reproducibile purpose
TrainingData3 <- createDataPartition(trainingProcessed2$classe, p=0.70, list=F)
TrainingData2 <- trainingProcessed2[TrainingData3, ]
TestingData2 <- trainingProcessed2[-TrainingData3, ]
```

## Prediction

### Random Forrest

Using the Random Forest model to predict activity
```{r, cache = T}
RandomForrest1 <- trainControl(method="cv", 5)
RandomForrest2 <- train(classe ~ ., data=TrainingData2, method="rf", trControl=RandomForrest1, ntree=250)
RandomForrest2
```

### Performance estimation
Estimation of performance against Validation data.  

```{r, cache = T}
RandomForrest3 <- predict(RandomForrest2, TestingData2)
confusionMatrix(TestingData2$classe, RandomForrest3)
```
```{r, cache = T}
RandomForrestAccuracy <- postResample(RandomForrest3, TestingData2$classe)
RandomForrestAccuracy
RandomForrestError <- 1 - as.numeric(confusionMatrix(TestingData2$classe, RandomForrest3)$overall[1])
RandomForrestError
```

### Estimation results
Estimated accuracy: 99.38%
Estimated out-of-sample error: 0.62%.


### Prediction against Test Data 
Using the prediction model against the Test data

```{r, cache = T}
RandomForrestPrediction <- predict(RandomForrest2, testingProcessed2[, -length(names(testingProcessed2))])
RandomForrestPrediction
```  

## Correlation Matrix
```{r, cache = T}
RandomForrestMatrix <- cor(TrainingData2[, -length(names(TrainingData2))])
corrplot(RandomForrestMatrix, method="circle")
```
